Determinism, Complexity, and Predictability in Computer Performance

### 2014-03-10 intro.tex
new abstract

[[Following is a good start at an abstract. It doesn't need to be
    reiterated in the introduction.]]  This paper provides insight
into when, why, and how forecast strategies fail when they are applied
to complicated time series by empirically estimating their inherent
complexity. [[Liz, do you think here would be better for the sentence
    about we are not the first, or is above where I asked Ryan to
    insert it ok? I worry it would muddy the water of this synopsis
    paragraph.]] We focus here on the specific example of computer
performance but believe the results apply to a much broader class of
time series. We conjecture that the complexity of traces from these
systems---which results from the inherent dimension, non-linearity,
and non-stationarity, as well as from measurement issues like noise,
aggregation, and finite data length---is \emph{empirically
  quantifiable}.  To validate this conjecture we model 120 different
performance traces using a variety of forecasting models.  We then
compare the empirically estimated \emph{complexity} with the accuracy
of each forecast method.

>>>>

%[[Following is a good start at an abstract. It doesn't %need to be
%    reiterated in the introduction.]]

### 2014-03-10 pecomplexity.tex
update macros to use xspace
new abstract

\documentclass[%
 reprint,
%superscriptaddress,
%groupedaddress,
%unsortedaddress,
%runinaddress,
%frontmatterverbose, 
%preprint,
%showpacs,preprintnumbers,
nofootinbib,
nobibnotes,
%bibnotes,
 amsmath,amssymb,
% aps,
%pra,
%prb,
%rmp,
%prstab,
%prstper,
%floatfix,
]{revtex4-1}
%\documentclass{article}
\usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
%\geometry{letterpaper}                   % ... or a4paper or a5paper or ...
%\geometry{landscape}                % Activate for for rotated page geometry
%\usepackage[parfill]{parIEEEabrvskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
%\usepackage{amssymb}
%\usepackage{amsmath}
\usepackage{epstopdf}
\usepackage{amsthm}

\usepackage{pifont} % check and x marks
\usepackage{color}

@@

\newcommand{\gcc}{{\tt 403.gcc}~}
\newcommand{\svd}{{\tt dgesdd}~}
\newcommand{\svdone}{{\tt dgesdd}$_1$~}
\newcommand{\svdtwo}{{\tt dgesdd$_2$}~}
\newcommand{\svdthree}{{\tt dgesdd$_3$}~}
\newcommand{\svdfour}{{\tt dgesdd$_4$}~}
\newcommand{\svdfive}{{\tt dgesdd$_5$}~}
\newcommand{\svdsix}{{\tt dgesdd$_6$}~}
\newcommand{\naive}{na\"ive}
\newcommand{\row}{{\tt row\_major}~}
\newcommand{\col}{{\tt col\_major}~}
%\newcommand{\color{red}}{{\color[rgb]{1,0,0}}}

@@

\begin{abstract}
{\color{green}From the site:It should be about 5\% of the length of the article, but less than 500 words.}[[Joshua: Needs to be rewritten to be more general and use computers as the application not the focus, should be done after we have a first draft of the paper]]  Computers are deterministic dynamical systems \cite{mytkowicz09}.
  Among other things, that implies that one should be able to use
  deterministic forecast rules to predict aspects of their behavior.
  That statement is sometimes---but not always---true. The memory and
  processor loads of some simple programs are easy to predict, for
  example, but those of more-complex programs like \gcc are not.
%%%%%%%%%%%%%%%%%%%%%%
%% I had to change all the \verb|blah| entries because they caused
%% latex to barf if they were in figure captions.  Odd bug.
%%%%%%%%%%%%%%%%%%%%%%
  The goal of this paper is to determine why that is the case. We
  conjecture that, in practice, complexity can effectively overwhelm
  the predictive power of deterministic forecast models. To explore
  that, we build models of a number of performance traces from
  different programs running on different Intel-based computers. We
  then calculate the \emph{permutation entropy}---a temporal entropy
  metric that uses ordinal analysis---of those traces and correlate
  those values against the prediction success.
\begin{description}
\item[Usage]
Secondary publications and information retrieval purposes.
\item[PACS numbers]
May be entered using the \verb+\pacs{#1}+ \pacs{05.45.Tp} command.
Time series analysis (nonlinear dynamics) 05.45.Tp
Information theory, 89.70.-a
	channel capacity in, 89.70.Kn
	communication complexity in, 89.70.Hj
	computational complexity in, 89.70.Eg
	entropy in, 89.70.Cf
	general biological information, 87.10.Vg
	in neuroscience, 87.19.lo

>>>>

\documentclass[%
 reprint,
%superscriptaddress,
%groupedaddress,
%unsortedaddress,
%runinaddress,
%frontmatterverbose, 
%preprint,
%showpacs,preprintnumbers,
nofootinbib,
nobibnotes,
%bibnotes,
 amsmath,amssymb,
% aps,
%pra,
%prb,
%rmp,
%prstab,
%prstper,
%floatfix,
]{revtex4-1}
%\documentclass{article}
\usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
%\geometry{letterpaper}                   % ... or a4paper or a5paper or ...
%\geometry{landscape}                % Activate for for rotated page geometry
%\usepackage[parfill]{parIEEEabrvskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
%\usepackage{amssymb}
%\usepackage{amsmath}
\usepackage{epstopdf}
\usepackage{amsthm}
\usepackage{xspace}
\usepackage{pifont} % check and x marks
\usepackage{color}

@@

\newcommand{\gcc}{{\tt 403.gcc}\xspace}
\newcommand{\svd}{{\tt dgesdd}\xspace}
\newcommand{\svdone}{{\tt dgesdd}$_1$\xspace}
\newcommand{\svdtwo}{{\tt dgesdd$_2$}\xspace}
\newcommand{\svdthree}{{\tt dgesdd$_3$}\xspace}
\newcommand{\svdfour}{{\tt dgesdd$_4$}\xspace}
\newcommand{\svdfive}{{\tt dgesdd$_5$}\xspace}
\newcommand{\svdsix}{{\tt dgesdd$_6$}\xspace}
\newcommand{\naive}{na\"ive}
\newcommand{\row}{{\tt row\_major}\xspace}
\newcommand{\col}{{\tt col\_major}\xspace}
%\newcommand{\color{red}}{{\color[rgb]{1,0,0}}}

@@

\begin{abstract}
{\color{green}From the site:It should be about 5\% of the length of the article, but less than 500 words.} This paper provides insight
into when, why, and how forecast strategies fail when they are applied
to complicated time series by empirically estimating their inherent
complexity. We focus here on the specific example of computer
performance but believe the results apply to a much broader class of
time series. We conjecture that the complexity of traces from these
systems---which results from the inherent dimension, non-linearity,
and non-stationarity, as well as from measurement issues like noise,
aggregation, and finite data length---is \emph{empirically
  quantifiable}.  To validate this conjecture we model 120 different
performance traces using a variety of forecasting models.  We then
compare the empirically estimated \emph{complexity} with the accuracy
of each forecast method.
\begin{description}
\item[Usage]
Secondary publications and information retrieval purposes.
\item[PACS numbers]
May be entered using the \verb+\pacs{#1}+ \pacs{05.45.Tp} command.
Time series analysis (nonlinear dynamics) 05.45.Tp
Information theory, 89.70.-a
	channel capacity in, 89.70.Kn
	communication complexity in, 89.70.Hj
	computational complexity in, 89.70.Eg
	entropy in, 89.70.Cf
	general biological information, 87.10.Vg
	in neuroscience, 87.19.lo
